Factors affecting health sector expenditure in OECD countries

### 2025-03-14 code/analysis-code/statistical-analysis.qmd
project #3

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      
library(kableExtra) 
library(webshot)
library(magick)
library(here)
library(modelsummary)
library(plm)
library(lmtest)
```

@@

# Save the regression results to an HTML file
modelsummary(
  model,
  output = here("results", "tables", "preliminary_regression_results.html"),
  title = "First Attempt at Regression Model",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  ),
  encoding = "UTF-8"  # Ensure UTF-8 encoding
)
```

@@

```{r}
# Perform Durbin-Watson test
dw_test <- dwtest(model)
print(dw_test)
```

@@

dw_test_lagged <- dwtest(fixed_model_lm)
print(dw_test_lagged)
```

>>>>

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      
library(kableExtra) 
library(webshot)
library(magick)
library(here)
library(modelsummary)
library(plm)
library(lmtest)
library(gt)
library(webshot2)

@@

modelsummary(
  model,
  output = here("results", "tables", "preliminary_regression_results.png"),
  title = "This table represents the preliminary regresssion model",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

@@

# Perform the Durbin-Watson test
dw_test <- dwtest(model)

# Extract the test statistic and p-value
dw_statistic <- dw_test$statistic
dw_p_value <- dw_test$p.value

# Create a data frame with the results
dw_results <- data.frame(
  Statistic = dw_statistic,
  P_Value = dw_p_value
)

# Generate a table using gt
dw_table <- gt(dw_results) %>%
  tab_header(
    title = "Durbin-Watson Test Results"
  ) %>%
  fmt_number(
    columns = vars(Statistic, P_Value),
    decimals = 4
  )

# Define the file path for the PNG
png_file <- here::here("results", "figures", "durbin_watson_test_results.png")

# Save the table as a PNG file
gtsave(dw_table, png_file)

@@

dw_test_fe <- dwtest(fixed_model_lm)
print(dw_test_fe)

# Extract the test statistic and p-value
dw_lstatistic <- dw_test_fe$statistic
dw_lp_value <- dw_test_fe$p.value

# Create a data frame with the results
dw_lresults <- data.frame(
  Statistic = dw_lstatistic,
  P_Value = dw_lp_value
)

# Generate a table using gt
dw_table <- gt(dw_lresults) %>%
  tab_header(
    title = "Durbin-Watson Test Results for FE "
  ) %>%
  fmt_number(
    columns = vars(Statistic, P_Value),
    decimals = 4
  )

# Define the file path for the PNG
png_file <- here::here("results", "figures", "durbin_watson_test_results_fe.png")

# Save the table as a PNG file
gtsave(dw_table, png_file)

### 2025-03-14 code/eda-code/eda.qmd
project #3

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(htmltools)
library(magrittr)
library(webshot2)
```

@@

>>>>

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(htmltools)
library(magrittr)
library(webshot2)
library(patchwork)
```

@@

# EDA (part2)

This is the EDA analysis performed for part 3 of the project.

```{r}
# Define dataset
df <- data_2010

# Function to create a scatter plot with correlation
create_corr_plot <- function(x_var, df, point_color) {
  cor_value <- round(cor(df[[x_var]], df$Current_health_expenditure_per_capita_current_US, use = "complete.obs"), 2)

  ggplot(df, aes_string(x = x_var, y = "Current_health_expenditure_per_capita_current_US")) +
    geom_point(color = point_color, alpha = 0.6) +  # Set color based on input
    geom_smooth(method = "lm", color = "red", se = FALSE) +
    labs(title = paste("Correlation:", cor_value), x = x_var, y = "Health Expenditure per Capita") +
    theme_minimal()
}

# Generate individual plots with different colors
plot1 <- create_corr_plot("GDP_growth_annual_", df, "blue")
plot2 <- create_corr_plot("Income_share_held_by_lowest_20", df, "blue")
plot3 <- create_corr_plot("Control_of_Corruption_Estimate", df, "blue")
plot4 <- create_corr_plot("Life_expectancy_at_birth_total_years", df, "blue")

# Arrange plots in a 2x2 grid
final_plot1 <- (plot1 + plot2) / (plot3 + plot4)
png_file1 <- here("results", "figures", "correlation_plot1.png")
ggsave(filename = png_file1, plot = final_plot1, width = 10, height = 8, dpi = 300)

# Generate individual plots with different colors
plot5 <- create_corr_plot("Literacy_rate_adult_total__of_people_ages_15_and_above", df, "green")
plot6 <- create_corr_plot("Trade__of_GDP", df, "green")
plot7 <- create_corr_plot("Foreign_direct_investment_net_inflows_BoP_current_US", df, "green")
plot8 <- create_corr_plot("Age_dependency_ratio__of_workingage_population", df, "green")

# Arrange plots in a 2x2 grid
final_plot2 <- (plot5 + plot6) / (plot7 + plot8)
png_file2 <- here("results", "figures", "correlation_plot2.png")
ggsave(filename = png_file2, plot = final_plot2, width = 10, height = 8, dpi = 300)
```

### 2025-03-14 products/manuscript/Manuscript.qmd
project #3

We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure. Some research have also focused on the relation between life expectancy and health expenditure. However, these research were based on fixed effect and random effect models. In our analysis we will try to follow OLS multiple regression model.

>>>>

We are using data from OECD members for analysis, and there have already been some research which have focused on GDP as predictor for health sector expenditure. Some research have such as Nkemgha in 2020 [@Nkemgha2020] studied the link among public health expenditure, private health expenditure and life expectancy in Cameroon. Christopoulos and Eleftheriou [@CHRISTOPOULOS2020195] using panel data analyzed the fiscal effects of health expenditures on health outcomes and found a significant impact of healthcare expenditures on increasing revenue. In our analysis we will try to follow OLS multiple regression model.
